\section{Discussion}
\label{sec:discussion}

\subsection{Non-Interchangeability of Outcome Statuses}
\label{sec:disc_noninterchangeability}

The central finding is that the four outcome statuses are not spatially interchangeable.
Jaccard coefficients below 0.15 for all pairs involving Located Dead demonstrate that the municipalities where lethal outcomes concentrate are fundamentally different from those where active disappearances or living recoveries cluster.
Even the highest pairwise overlap, Total and Located Alive at $J = 0.674$, leaves one-third of hotspot municipalities unshared.

This result parallels two established findings in spatial criminology.
\citet{andresen_2006} showed that counts and rates of the same crime type produce different hotspot geographies; \citet{braga_law_2017} argued that crime-type disaggregation matters for concentration analysis.
Our finding operates on a third axis: disaggregation by \emph{resolution category} within a single crime type produces non-overlapping spatial clusters.

Forensic search and prevention programs require distinct geographic targets. Forensic teams prioritize recovery sites, while prevention programs must focus on where people vanish. Because composite maps are 60\% Located Alive cases, they misallocate resources by masking the municipalities where lethal outcomes concentrate.

\subsection{Regional Dynamics}
\label{sec:disc_regional}

Centro's emergence as the fastest-growing HH region is the most relevant temporal finding, with Not Located expanding at $+4.96$ municipalities per year ($p < 0.001$).
Prior literature on Mexico's disappearance crisis has focused on the Norte and Baj\'{i}o axis, where cartel territorial competition is well documented \citep{cadena_garrocho_2019, rios_2013}.
The Centro expansion went undetected in annual analyses because it began around 2019 and accelerated through 2024 and 2025, a trajectory visible only at monthly resolution.
Fourth competing hypotheses may explain this pattern, and we cannot distinguish between them with the available data.
The first is a genuine increase in disappearances across the metropolitan corridor spanning Estado de M\'{e}xico, CDMX, Puebla, Morelos, and Hidalgo, potentially reflecting the urbanization of organized criminal violence as trafficking routes shift toward the capital region \citep{vilalta_muggah_2016, medel_2015}.
The second is a reporting artifact: urban municipalities with stronger prosecutorial capacity and more active civil society may register a higher proportion of existing cases without a corresponding increase in underlying incidence \citep{heredia_gonzalez_2022}.
The third is geographic displacement: enforcement pressure in Norte and Baj\'{i}o may have pushed criminal operations into Centro, consistent with the ``cockroach effect'' documented in Latin American security studies \citep{dell_2015, guerrero_gutierrez_2011, bagley_2013}. Last, the institutional reforms documented by the Mexican government since 2019: these include the establishment of state-level search commissions and expanded RNPDNO coverage, coincide with the largest expansion of Not Located hotspot municipalities observed in our study. This temporal coincidence suggests that legislative and administrative measures have not yet translated into a measurable spatial contraction of the disappearance crisis. While this trend represents a failure to contain the territorial growth of violence, it may also partly reflect improved registration capturing previously invisible cases.


Norte's consistent positive slopes across all four statuses ($p < 0.05$ for each) suggest an entrenched hotspot that is simultaneously deepening and revealing its outcomes.
The coexistence of growing Located Alive and Located Dead HH clusters may reflect a combination of high organized crime incidence and relatively developed forensic infrastructure in states such as Tamaulipas and Nuevo Le\'{o}n \citep{atuesta_connection_2025}.
These border municipalities function not only as sites of local violence but as nodes in transnational trafficking networks, where the intersection of migration corridors, drug supply chains, and cross-border financial flows concentrates multiple forms of criminal activity \citep{reyes_remesas_2023}.
This interpretation is a hypothesis; it could alternatively reflect that Norte's crisis has reached a scale at which all outcome categories produce detectable clustering.

% REWRITTEN: Bajío paragraph using 29-municipality corridor numbers
The ``rising Baj\'{i}o'' hypothesis, suggested by \citeauthor{cadena_garrocho_2019}'s (\citeyear{cadena_garrocho_2019}) 2006 to 2017 analysis and attributed to cartel route reorganization through the Guanajuato--Jalisco corridor, finds qualified support in the post-2017 period when analyzed at the 29-municipality industrial corridor scale.
A Chow test identifies a structural break at February 2017 ($F = 78.13$, $p < 0.001$).
Prior to this break, HH municipalities within the corridor were declining ($\beta_1 = -8.82$/yr); after 2017, the corridor shows a slow but statistically significant expansion ($\beta_2 = +0.47$/yr, $p = 0.046$), driven primarily by Not Located cases ($+0.50$/yr, $p < 0.001$).
The corridor is gaining places where people disappear and are not found, even as the overall number of hotspots remains broadly stable ($-0.31$/yr, $p = 0.041$), reflecting the net balance of a growing Not Located footprint against a shrinking Located Alive footprint ($-0.35$/yr, $p = 0.003$).
This pattern is consistent with the intensification of cartel competition in Guanajuato's industrial municipalities, particularly around fuel infrastructure and territorial control \citep{le_cour_2020}.
The February 2017 breakpoint precedes the November 2017 General Law by eight months; institutional anticipation effects and state-level pilot programs cannot be ruled out as contributing factors.

\subsection{CED Alignment}
\label{sec:disc_ced}

Our analysis provides the first national-scale quantitative evidence bearing on the CED's characterization of disappearances in Mexico as occurring at multiple moments and across different parts of the territory \citep{CED_MEX_art34_2026}.
The LISA results show HH clusters emerging, expanding, and shifting across states over \nmonths{} months, a pattern consistent with the CED's temporal and geographic narrative.
Three qualifications apply.
First, the RNPDNO aggregates all disappearances regardless of perpetrator; our clusters capture the composite phenomenon, not enforced disappearances specifically.
The Mexican government has argued that many registered cases reflect private-actor violence \citep{CED_MEX_visit_2022}; our analysis is agnostic to this distinction.
Second, several CED-named states (Coahuila, Veracruz, Nayarit) show low or zero HH municipalities in our study window, which may reflect underreporting rather than absence of crisis, as their most acute periods (2009 to 2017) largely precede our data.
% REWRITTEN: Jalisco paragraph now volume-focused, not spatial-diffusion-focused
Third, the Jalisco result requires careful interpretation given the CED's characterization of the state as having the highest current disappearance count.
RNPDNO registrations in Jalisco collapsed from 3,825 persons in 2021 to 216 in 2024, a 94\% decline, with a corresponding drop in municipalities reporting cases (from 103 to 41).
The absence of HH clusters in recent years therefore reflects a registration-volume collapse rather than a spatial dispersion of violence.
Whether this represents a genuine decline in disappearances, a shift in reporting practices, or a reclassification of cases under other legal categories cannot be determined from the registry alone, but the discrepancy between RNPDNO trajectories and the CED's characterization warrants independent verification.

\subsection{Sex Patterns}
\label{sec:disc_sex}

The Located Alive near-parity (49.4\% male, 50.5\% female) is striking given the male-dominated profile of Not Located cases (77.9\% male).
Women constitute a larger share of Located Alive registrations than of Not Located registrations, a compositional pattern that warrants investigation.
Two mechanisms are plausible: gender-specific disappearance types with different recovery profiles, such as domestic violence or trafficking cases that may resolve at higher rates \citep{frias_2023}, or differential registration practices in which women's recoveries are recorded more systematically due to advocacy pressure \citep{schwartz_marin_2016}.

The spatial divergence in Not Located male versus female HH (Jaccard = 0.262, with only 28 of 107 municipalities shared) indicates that these are geographically distinct phenomena, not simply different counts in the same places.
The female Located Dead power failure, zero HH municipalities consistent with a mean of 0.0065 cases per municipality-month and 99.4\% zero cells, underscores a general limitation of LISA applied to rare events.

\subsection{External Validation}
\label{sec:disc_vawri}

The VAWRI comparison confirms that HH municipalities coincide with areas of elevated recorded violence regardless of victim sex, with large effect sizes for both male and female classifications.
This convergence indicates that the LISA clusters capture genuine violence concentrations rather than statistical artifacts; both sexes cluster in high-violence municipalities.


\subsection{Limitations}
\label{sec:limitations}

\begin{enumerate}

  \item \textbf{Raw counts without population normalization.}
        Concentration in large, high-population municipalities may partly reflect population size rather than elevated risk.
        A population-normalized companion analysis is forthcoming \citep[cf.][]{andresen_2006}.

  \item \textbf{Ecological fallacy.}
        Municipal-level analysis cannot support individual-level inferences about victimization risk or perpetrator behavior.

  \item \textbf{Missing data.}
        Municipality-months absent from the RNPDNO are treated as zero cases.
        If absences reflect reporting failure rather than true zeros, Gini and HHI are overestimated (appearing more concentrated than reality), while the number of HH municipalities is underestimated (the true geographic spread is broader).
        More generally, disappearances in municipalities with weak state presence or low civil society activity may be systematically underreported, biasing hotspot maps toward high-capacity jurisdictions \citep{guercke_2021}.

  \item \textbf{Spatial weights specification.}
        Queen contiguity may not capture functional connectivity such as road networks or cartel territorial boundaries.
        The fuzzy\_contiguity specification ($50$\,m buffer, zero islands) mitigates the most common boundary-alignment failures, but sensitivity to alternative weight matrices (k-nearest neighbors, distance-based) is left for future work.

  \item \textbf{Descriptive design.}
        This analysis documents spatial patterns; it does not identify causal mechanisms.
        All regional narratives are hypotheses requiring econometric testing with external covariates.
        Female Located Dead produces zero HH municipalities due to extremely low base rates (mean 0.0065 per municipality-month), and LISA results for this subgroup should not guide policy.

  \item \textbf{Perpetrator heterogeneity.}
        The RNPDNO does not distinguish enforced disappearances (state involvement) from those by private actors; our clusters capture both.
        This limits the interpretability of hotspots as indicators of any single causal process.

  \item \textbf{End-of-period reporting lag.}
        November to December 2025 totals show ratios of 0.46 to 0.72 relative to the same status's July 2025 count,\footnote{Ratios computed as Dec 2025 national total divided by Jul 2025 national total for each status, indicating incomplete registration in the most recent months.} indicating incomplete registration.
        June cross-sections mitigate this for snapshot analyses, but full-series trend estimates include these partially reported months.

  \item \textbf{Cross-section sensitivity.}
        Snapshot analyses use June of each year.
        Results may differ under alternative month selections; sensitivity to cross-section choice is not formally tested.

  \item \textbf{Registration geography versus event geography.}
        The RNPDNO records the municipality of last known contact, not the site of the criminal act.
        For Located Dead, this may differ from the municipality where remains were recovered, introducing a geographic displacement between the registered disappearance and the lethal outcome.
        The Jaccard divergence therefore captures differences in the geography of registration rather than necessarily distinct causal processes.

\end{enumerate}