\section{Methods}
\label{sec:methods}

\subsection{Spatial Weights}
\label{sec:weights}
We construct a row-standardized Queen contiguity weights matrix for all \nmunis{} municipalities.
To correct sub-metre boundary misalignments in the INEGI cartographic source, adjacency is determined using a 50-metre buffer tolerance rather than exact topological intersection, eliminating spurious island municipalities that would otherwise be excluded from spatial autocorrelation analysis.
This yields \nislands{} true island municipalities.

\subsection{Global Spatial Autocorrelation}
\label{sec:global}

Global Moran's $I$ quantifies the overall degree of spatial clustering:

\begin{equation}
  I = \frac{n}{\sum_i \sum_j w_{ij}}
      \cdot
      \frac{\sum_i \sum_j w_{ij}(x_i - \bar{x})(x_j - \bar{x})}
           {\sum_i (x_i - \bar{x})^2}
  \label{eq:moran}
\end{equation}

\noindent where $x_i$ is the case count in municipality $i$, $w_{ij}$ are row-standardized Queen contiguity weights, and $n = \text{\nmunis{}}$.
Statistical significance is assessed via 999 conditional permutations at $\alpha = 0.05$.

\subsection{Local Spatial Autocorrelation (LISA)}
\label{sec:lisa}
Local Moran's $I_i$ decomposes the global statistic into municipality-level contributions \citep{anselin_1995}:
\begin{equation}
  I_i = \frac{(x_i - \bar{x})}{m_2} \sum_j w_{ij}(x_j - \bar{x}), \quad
  m_2 = \frac{1}{n}\sum_i (x_i - \bar{x})^2
  \label{eq:lisa}
\end{equation}
\noindent Municipalities are classified into five types based on the sign of their local statistic and the significance of their pseudo-$p$-value at $\alpha = 0.05$: High-High (HH), Low-Low (LL), High-Low (HL), Low-High (LH), and Not Significant (NS).
HH municipalities have high counts surrounded by high-count neighbors; LL municipalities have low counts surrounded by low-count neighbors.

We extract connected-component HH clusters via breadth-first search on the Queen adjacency graph, retaining only clusters of three or more contiguous HH municipalities.
The minimum-three threshold excludes isolated municipality pairs that may achieve HH significance through boundary artifacts without forming a spatially coherent cluster.
This spatial filter serves as the primary guard against false positives.

No family-wise or false discovery rate correction is applied to the individual LISA $p$-values.
The conditional permutation approach is already conservative relative to parametric inference, and the BFS filter provides an additional structural constraint.
We assessed Benjamini-Hochberg FDR correction at $q = 0.05$; however, with 999 permutations the minimum achievable pseudo-$p$ is 0.001, while the BH-adjusted threshold for \nmunis{} simultaneous tests falls below $2 \times 10^{-5}$, a permutation resolution floor that eliminates nearly all classifications regardless of their substantive validity.
We therefore retain uncorrected $\alpha = 0.05$ with the spatial BFS filter as the primary threshold and report sensitivity at $\alpha = 0.01$ in Appendix~A.2, where 33 to 49\% of HH municipalities survive with the Jaccard structure preserved.

LL clusters are retained in tables for completeness but are not interpreted as a substantive finding: the extreme zero-inflation of the count data (77 to 98\% of municipality-months record zero cases, depending on status) means that LL classification largely identifies municipalities where nothing happened surrounded by municipalities where nothing happened.

The total computational scope is \nLISAruns{} LISA runs: 4 statuses $\times$ 3 sex categories (total, male, female) $\times$ \nmonths{} months, each computed over $N = \text{\nmunis{}}$ municipalities with 999 permutations.
As a temporal stability check, we compute Spearman rank correlations of local $I_i$ values between consecutive June cross-sections.

\subsection{Geographic Concentration}
\label{sec:concentration}

We quantify geographic concentration using the Gini coefficient:

\begin{equation}
  G = \frac{2\sum_{i=1}^{n} i \cdot x_{(i)}}{n \sum_{i=1}^{n} x_{(i)}}
      - \frac{n+1}{n}
  \label{eq:gini}
\end{equation}

\noindent where $x_{(i)}$ is the $i$-th order statistic (sorted ascending) of the count distribution, and the Herfindahl-Hirschman Index (HHI):

\begin{equation}
  \text{HHI} = \sum_{i=1}^{n} s_i^2, \quad s_i = \frac{x_i}{\sum_j x_j}
  \label{eq:hhi}
\end{equation}

\noindent where $x_i$ is the raw (unsorted) count for municipality $i$.
Both metrics are computed for each of the \nmonths{} monthly cross-sections per status.
We benchmark our concentration estimates against \citet{weisburd_2015}'s ``law of crime concentration,'' which posits that approximately 50\% of crime concentrates in 4--6\% of places.

\subsection{Trend Analysis}
\label{sec:trend}

We fit OLS regressions of the form $y_t = \alpha + \beta t + \varepsilon_t$ separately for each combination of six regions and four statuses, where $y_t$ is the monthly count of HH municipalities and $t = 1, \ldots, \text{\nmonths{}}$.
Five regions follow the CESOP classification \citep{mexicoencifras}: Norte, Norte-Occidente, Centro-Norte, Centro, and Sur.
The sixth unit is the 29-municipality Baj\'{i}o industrial corridor. Slopes are expressed as municipalities per year ($\hat\beta \times 12$).
Because monthly HH counts exhibit serial correlation, standard errors are computed using the \citet{newey_west_1987} heteroskedasticity and autocorrelation consistent (HAC) estimator with a bandwidth of 12 months, corresponding to one year of seasonal periodicity.

% CHANGED: Bajío breakpoint test method
For the Baj\'{i}o corridor, where visual inspection suggests a non-monotonic trajectory, we additionally test for structural breaks using the Chow test at monthly candidate breakpoints from January 2016 through December 2021.
Where a break is confirmed, we fit a piecewise linear model with Newey-West HAC standard errors and report pre- and post-break slopes separately.

To test whether the regional distribution of HH municipalities changed over the study period, we compute Pearson chi-square statistics comparing the observed distribution of HH municipalities across regions in \bookendstart{} against the distribution in \referencemonth{}, separately for each status.
We use June of each year as the reference cross-section to avoid end-of-year reporting lags documented in Section~\ref{sec:limitations}.

\subsection{Sex Disaggregation}
\label{sec:sex_methods}

LISA is repeated separately for male and female case counts across all statuses and months.
We compare male and female HH geographies using Jaccard similarity for each status at the \referencemonth{} cross-section.
A critical limitation applies to female Located Dead: with a mean of 0.0065 cases per municipality-month and 99.4\% zero cells, the LISA algorithm lacks sufficient spatial variation to detect meaningful clusters.
Female Located Dead results are reported for completeness but should be interpreted with  caution.



\subsection{External Validation}
\label{sec:vawri_methods}
As a plausibility check on the HH classification, we compare Violence Against Women Risk Index (VAWRI) scores \citep{escobar_vawri_2026}, a composite measure of reported gender-based violence at the municipal level, in HH versus non-HH municipalities at the June 2023 cross-section, matching the VAWRI reference year.
Comparisons are conducted separately for each status and, for Not Located and Located Dead, separately by sex.
Differences are tested using the Mann-Whitney $U$ test with rank-biserial correlation as effect size \citep{kerby_2014}, appropriate given the skewed distribution of VAWRI scores across municipalities.

\subsection{CED Alignment Analysis}
\label{sec:ced_methods}

To situate our spatial findings within the policy landscape, we extract HH dynamics for eight states explicitly named in the Committee on Enforced Disappearances' Article~34 decision \citep{CED_MEX_art34_2026}: Jalisco, Guanajuato, Coahuila, Veracruz, Nayarit, Tabasco, Nuevo Le\'{o}n, and Estado de M\'{e}xico.
For each state, we count the number of municipalities classified as HH in each month, overlay the CED's temporal claims, and compute OLS trend slopes.
%No additional LISA runs are required; this analysis operates entirely on the existing \nLISAruns{} runs.